---
title: "Homework 2"
author: "Nancy Yeung (ny3257) - SDS 315"
output: pdf_document
---
\vspace{-1cm}
```{r global_options, echo=FALSE}
knitr::opts_chunk$set(fig.height=4, fig.width=9, warning=FALSE, message=FALSE, echo=FALSE)
```

```{r}
library(ggplot2)
library(RColorBrewer)
library(dplyr)
library(kableExtra)
library(knitr)
library(tidyverse)
Profs <- read.csv('profs.csv')
Bikeshare <- read.csv('bikeshare.csv')
Capmetro <- read.csv('capmetro_UT.csv')
Billboard <- read.csv('billboard.csv')
```
------------------------------------------------------------------------
\begingroup
\fontsize{12}{14}\selectfont
**Problem 1: Beauty, Or Not, in the Classroom**
\endgroup

\begingroup
\fontsize{10}{12}\selectfont
**Part A**
\endgroup

The histogram below illustrates the distribution of overall course evaluation scores. The x-axis represents the scores, while the y-axis indicates the frequency at which each score was given.

```{r}
#Histogram of the distribution of overall course evaluation scores
ggplot(Profs) + geom_histogram(aes(x=eval), fill = 'skyblue', color = 'white') + labs(title = 'Distribution of Course Evaluation Scores', x = 'Course Evaluation', y = 'Frequency', caption = paste("The distribution of the graph is slightly left-skewed, with a median and range of", median(Profs$eval), "and", max(Profs$eval)-min(Profs$eval), "respectively. \nThis indicates that, based on the sampled courses, students generally provide positive ratings for their professors.")) + theme_minimal() + theme(text=element_text(size = 13, family="serif"))
```

\begingroup
\fontsize{10}{12}\selectfont
**Part B**
\endgroup

The below boxplots show the distribution of course evaluation scores split by whether or not the professor is a native English speaker. 

```{r}
#Box plots of the distribution of overal course evaluation scores split by whether the professor is a na
ggplot(Profs) + geom_boxplot(aes(x=eval, fill=native)) + labs(title = 'Distribution of Course Evaluation Scores by Fluency', x = 'Course Evaluation', x = NULL, fill = 'Native English Speaker', caption = paste("The median score for course evaluations is higher for professors who are native speakers, however, there were several \noutliers in the evaluations for native speakers, whereas there were no outliers for non-native speakers. Overall, native \nEnglish speaking professors received higher ratings, consequently, they also received some of the lowest scores.")) + scale_fill_manual(values = c('pink', 'skyblue')) + theme_minimal() + coord_flip() + theme(text=element_text(size = 13, family="serif"), axis.text.x = element_blank())
```

\begingroup
\fontsize{10}{12}\selectfont
**Part C**
\endgroup

The histograms below compares the distribution of course evaluation scores for male and female professors.

```{r}
#Histograms of the distribution of course evaluation scores for male and female professors
ggplot(Profs) + geom_histogram(aes(x=eval, fill=gender), color = 'white') + facet_wrap(~gender) + labs(title = 'Distribution of Course Evaluation Scores by Gender', x = 'Course Evaluation', y = 'Frequency', fill = 'Gender', caption = paste("Both of the graphs are left-skewed. Male professors had overall higher evaluation \nscores compared to female professors with the medians being", Profs %>% filter(gender == 'male') %>% summarise(median_eval = median(eval)) %>% pull(median_eval), "and", Profs %>% filter(gender == 'female') %>% summarise(median_eval = median(eval)) %>% pull(median_eval), "respectively.")) + scale_fill_manual(values = c('pink', 'skyblue')) + theme_minimal() + theme(text=element_text(size = 13, family="serif"), strip.text = element_blank())
```

\begingroup
\fontsize{10}{12}\selectfont
**Part D**
\endgroup

The scatterplot below shows the extent of the association between the professor's physical attraction and the course evaluation score they received.

```{r}
#Scatterplot of the association between the professor's physical attraction and the course evaluation score they received
ggplot(Profs) + geom_point(aes(x=beauty, y=eval), color = 'skyblue') + labs(title = 'Association Between Looks and Course Evaluation', x = 'Attractiveness', y  = 'Course Evaluation', caption = paste("The correlation between a professor's physical attractiveness and course evaluation score is", round(cor(Profs$beauty, Profs$eval), 2),". Based on both \nthe correlation coefficient as well as the distribution of points on the graph, there is a very small correlation between these two variables.")) + theme_minimal() + theme(text=element_text(size = 13, family="serif"))
```
\newpage

\begingroup
\fontsize{12}{14}\selectfont
**Problem 2: Bike Sharing**
\endgroup

```{r}
#Creating a new dataset AvgRentals that has the average hourly bike rentals
AvgRentals <- Bikeshare %>%
  group_by(hr) %>%
  summarise(AvgTotal = mean(total))
AvgRentals <- merge(AvgRentals, Bikeshare)
```

\begingroup
\fontsize{10}{12}\selectfont
**Part A**
\endgroup

The line graph below shows the average hourly bike rentals across all hours of the day.

```{r}
#Line plot of the average hourly bike rentals across all hours of the day
ggplot(AvgRentals) + geom_line(aes(x=hr, y=AvgTotal), color = 'skyblue') + labs(title = 'Average Total Bike Rentals Across Different Hours of the Day', x = 'Hour', y = 'Average Total Bike Rentals', caption = paste("Bike rentals reached a peak at hour", AvgRentals$hr[which.max(AvgRentals$AvgTotal)], "with an average of", round(max(AvgRentals$AvgTotal), 2), "bikes rented. \nPrior to hour 5, the average bike rentals were relatively low, gradually increasing and reaching a peak before experiencing a subsequent decline.")) + theme_minimal() + theme(text=element_text(size = 13, family="serif"))
```

\begingroup
\fontsize{10}{12}\selectfont
**Part B**
\endgroup

A faceted line graph showing average bike rentals by hour of the day, split according to whether it is a working day or not.

```{r}
#Detecting if it is a working day or not and labeling them appropriately 
AvgRentals$type <- ifelse(AvgRentals$workingday == 1, 'Workday', 'No Workday')

#Plotting a faceted line graph showing average bike rentals by hour of the day, split according to whether it is a working day or not
AvgRentals %>%
  group_by(type, hr) %>%
  summarise(AvgTotal = mean(total)) %>%
  ggplot() + geom_line(aes(x = hr, y = AvgTotal, color = type)) + labs(title = 'Average Total Bike Rentals Across Different Hours of the Day by Weekday', x = 'Hour', y = 'Average Total Bike Rentals', color = 'Day Type', caption = paste("Average bike rentals reach a higher peak on weekdays with over 500 rentals whereas on non workdays they peak at a little over 350. On \nworkdays, after peak hours, average bike rentals experience sharp declines compared to non-workdays where it's a slow, steady decline.")) + scale_color_manual(values = c('Workday' = 'skyblue', 'No Workday' = 'pink')) + theme_minimal() + facet_wrap(~type, scales = 'free_y') + theme(text=element_text(size = 13, family="serif"))
```

\newpage

\begingroup
\fontsize{10}{12}\selectfont
**Part C**
\endgroup

A faceted bar plot showing average ridership during 9 AM by weather situation code, faceted according to whether it is a working day or not. Weather code 1 corresponds to clear skies, few clouds, partly cloudy conditions, or a mix of clear and cloudy weather. Weather code 2 represents mist with cloudy conditions, mist with broken clouds, mist with a few clouds, or just mist. Weather code 3 signifies light snow, light rain with thunderstorms and scattered clouds, or light rain with scattered clouds. Weather code 4 denotes heavy rain with ice pellets and thunderstorms in misty conditions, or snow with foggy weather.

```{r}
#Filtering so only info for bike rentals at 9 AM are shown and then calculating the average
RentalsNine <- filter(Bikeshare, hr == 9)
AvgRentalsNine <- RentalsNine %>%
  group_by(weathersit, workingday) %>%
  summarise(AvgTotal = mean(total))

#Plotting a faceted bar graph showing average ridership during 9 AM by weather situation code, faceted according to whether it is a working day or not
ggplot(AvgRentalsNine) + geom_bar(aes(x = weathersit, y = AvgTotal, fill = factor(workingday)), stat = "identity") + facet_wrap(~workingday, labeller = labeller(workingday = c('0' = 'No', '1' = 'Yes'))) + labs(title = 'Average Ridership by Weather Situation at 9 AM by Weekday', x = 'Weather Situation', y = 'Total Bike Rentals', fill = 'Workingday', caption = paste("Average ridership at 9 AM is higher on workdays no matter what the wheather situation is. \nReguardless of what type of day it is, if there is light rain or snow average bike ridership will go down.")) + scale_fill_manual(name = 'Workingday', labels = c('No', 'Yes'), values = c('pink', 'skyblue')) + theme(legend.position = "none") + theme_minimal() + theme(text=element_text(size = 13, family="serif"))
```

\newpage

\begingroup
\fontsize{12}{14}\selectfont
**Problem 3 - Capital Metro UT Ridership**
\endgroup

```{r}
#Recode the categorical variables in sensible, rather than alphabetical, order
Capmetro = mutate(Capmetro,
day_of_week = factor(day_of_week,
levels=c("Mon", "Tue", "Wed","Thu", "Fri", "Sat", "Sun")),
month = factor(month,
levels=c("Sep", "Oct","Nov")))

#Calculating the average boarding
AvgCapmetro <- Capmetro %>%
  group_by(hour_of_day, day_of_week, month) %>%
  summarise(AvgBoarding = mean(boarding))
AvgCapmetro <- merge(AvgCapmetro, Capmetro)
```

\begingroup
\fontsize{10}{12}\selectfont
**Part A**
\endgroup

Below are faceted line graph that plot average boardings by hour of the day, day of week, and month. Each graph is a different day of the week and the three lines show data for the months September, October, and November, while the x-axis shows the hour of the day and the y shows the average boarding. 
```{r}
#Faceted line graph that plot average boardings by hour of the day, day of week, and month
ggplot(AvgCapmetro) + geom_line(aes(x=hour_of_day, y=AvgBoarding, color = factor(month))) + facet_wrap(~day_of_week) + labs(title = 'Average Boardings by Hour of the Day, Day of the Week, and Month', x = 'Hour of the Day', y = 'Average Boardings', color = 'Month', caption = paste("On weekdays peak boarding occurs at around the same time, while on weekends boarding numbers are consistently low throughout the day. \nLow average boarding on Mondays in September could be a result of students not being back from summer vacations yet. \nLow boardings Weds/Thurs/Fri in November could be for a similar reason with students returning home early for Thanksgiving break.")) + scale_color_manual(values = c('skyblue', 'violet', 'pink')) + theme(text=element_text(size = 13, family="serif"))
```

\newpage

\begingroup
\fontsize{10}{12}\selectfont
**Part B**
\endgroup
The faceted scatter plots below showing boardings vs. temperature, faceted by hour of the day, and with points colored in according to whether it is a weekday or weekend. 

```{r}
#Scatter plot showing boardings vs. temperature, faceted by hour of the day, and with points colored in according to whether it is a weekday or weekend
ggplot(AvgCapmetro) + geom_point(aes(x=temperature, y=boarding, color = factor(weekend)), size=.5) + facet_wrap(~hour_of_day) + labs(title = 'Boardings vs. Temperature by Hour of the Day', x = 'Temperature', y = 'Boardings', color = 'Type of Day', caption = paste("Disregarding the type of day it is, we can see that temperature has little to no effect on whether students ride the bus. \nStudents take the bus more often on weekdays.")) + scale_color_manual(values = c('pink', 'skyblue')) + theme(text=element_text(size = 13, family="serif"))
```

\newpage

\begingroup
\fontsize{12}{14}\selectfont
**Problem 4: Wrangling the Billboard Top 100**
\endgroup

\begingroup
\fontsize{10}{12}\selectfont
**Part A**
\endgroup

The table below shows the top 10 most popular songs since 1958, as measured by the total number of weeks that a song spent on the Billboard Top 100. Note that the data used ends in week 22 of 2021, so the most popular songs of 2021 from that period on wards will not have be recorded. The table has three columns: performer, song, and count, with count representing the number of weeks that song appeared in the Billboard Top 100. 

```{r}
#Arranging the dataset in descending order based on weeks spent on the chart
Billboard <- arrange(Billboard, desc(weeks_on_chart))

#Keeping only the first 10 results
Top10 <- Billboard %>%
  distinct(song, .keep_all = TRUE) %>%
  head(n = 10)

#Getting rid of unnecessary columns
BillboardTable <- subset(Top10, select = -c(X, url, week_id, week_position, song_id, instance, previous_week_position, year, week, peak_position))

#Creating the kable table
kbl(BillboardTable, booktabs = T, col.names = c("Song", "Performer", "Weeks on Chart")) %>%
  kable_styling(latex_options = c("striped", "hold_position"))
```

\begingroup
\fontsize{10}{12}\selectfont
**Part B**
\endgroup

The line graph below shows mucial diversity in the Billboard Top 100 from 1959 to 2020.

```{r}
#Filtering the data to only include the years 1959-2020
PartB <- filter(Billboard, year %in% seq(1959, 2020))

#Grouping the data by year and unique id, before counting
Songs <- PartB %>%
  group_by(year, song_id) %>% 
  summarise(count = n())

#Grouping by year and then finding the number of unique songs for each year
UniqueSongs <- Songs %>%
  group_by(year) %>%
  summarise(unique_songs = n())

#Creating a line plot of the musical diversity in the billboard 100 over this time period
ggplot(UniqueSongs) + geom_line(aes(x = year, y = unique_songs), color = 'skyblue') + labs(title = "Musical Diversity of Billboard Top 100 Over Time", x = "Year", y = "Number of Unique Songs on Top 100", caption = "Musical diversity peaked around 1965, before experiencing a sharp decline and hitting and all time low in the 2000's. \nHowever by 2020 musical diversity has almost rised back to its former peak.") + theme_minimal() + theme(text=element_text(size = 13, family="serif"))
```

\newpage

\begingroup
\fontsize{10}{12}\selectfont
**Part C**
\endgroup

The bar plot below shows 19 artists who have had at least 30 songs that appeared on the Billboard Top 100 for at least 10 weeks, also dubbed as a "10-week hit". The y-axis shows how many of these "10-week hits" each of these 19 artists have.

```{r}
WeekHit <- filter(Billboard, weeks_on_chart >= 10)
UniqueWeekHit <- WeekHit %>%
  distinct(song_id, .keep_all = TRUE)

UniqueWeekHit <- UniqueWeekHit %>%
  group_by(performer) %>%
  summarise(WeekHit10 = n())

TopPerformer <- UniqueWeekHit %>% 
  filter(WeekHit10 >= 30)

ggplot(TopPerformer, aes(x = reorder(performer, -WeekHit10), y = WeekHit10)) + geom_bar(stat = "identity", fill = "skyblue") + labs(title = "Top Artists with at Least 30 Ten-Week Hits on Billboard Top 100", x = "Artist", y = "Number of Ten-Week Hits", caption = "Elton John has had the most 10-week hits, consequently Brad Paisley, Daryl Hall John Oates, \nGeorge Strait, Jason Aldean, and Neil Diamond are all tied with 30 10-week hits for the least.") + theme_minimal() + theme(axis.text.x = element_text(angle = 45, hjust = 1)) + theme(text=element_text(size = 13, family="serif"))
```